`r paste(params$species)` biomass trends and figures

### 2019-10-21 analysis/VOCC/2-biomass-trends.Rmd
change years plotted for biomass trends

---
title: "`r paste(params$species)` biomass trends and figures"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    species: "Lingcod"
    region: "All synoptic surveys"
    bind_predictions: FALSE
    covs: "-trawled-ssid" # covariate string used to name model rds 
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
                      #fig.path=paste0("figs/", spp, "/"),
                      echo=FALSE, warning=FALSE, message=FALSE)
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)
```

```{r params}
species <- params$species
region <- params$region
bind_predictions <- params$bind_predictions
covs <- params$covs
priors <- FALSE

@@

if (region == "All synoptic surveys") {
  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
  model_ssid <- c(1, 3, 4, 16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "Both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}
```

@@

if (region != "All synoptic surveys") {
  try(adult_predictions <- readRDS(paste0("data/", spp,
    "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"
  )))
  try(imm_predictions <- readRDS(paste0("data/", spp,
    "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds"
  )))

@@

    try(predictions <- readRDS(paste0("data/", spp,
      "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"
    )))

@@

Caluclate variables of interest from predictions
```{r}
if (exists("adult_predictions")) {

  # to convert from kg/m2 to kg/hectare multiply by 10000
  adult_predictions$est_exp <- exp(adult_predictions$est) * 10000
  imm_predictions$est_exp <- exp(imm_predictions$est) * 10000

@@

@@

  model_ssid <- unique(adult_predictions$ssid)
} else {
  predictions$total_bio <- exp(predictions$est) * 10000
  max_raster <- quantile(predictions$total_bio, .999)
  max_bio <- round(max(predictions$total_bio))

@@

### Split by survey (SSID)
```{r}
if (exists("adult_predictions")) {
  if (1 %in% model_ssid) {
    adult_predictions1 <- adult_predictions %>% filter(ssid == 1)
  } else {
    adult_predictions1 <- NULL
  }

  if (3 %in% model_ssid) {
    adult_predictions3 <- adult_predictions %>%
      filter(ssid != 16) %>%
      filter(ssid != 4)
  } else {
    adult_predictions3 <- NULL
  }

  if (4 %in% model_ssid) {
    adult_predictions4 <- adult_predictions %>% filter(ssid == 4)
  } else {
    adult_predictions4 <- NULL
  }

  if (16 %in% model_ssid) {
    if (any(adult_predictions$ssid == 16)) {
      adult_predictions16 <- adult_predictions %>% filter(ssid == 16)
    } else {
      adult_predictions16 <- NULL
    }
  } else {
    adult_predictions16 <- NULL
  }

@@

  p_adult <- list()
  for (i in seq_len(length(model_ssid))) {
    p_adult[[i]] <- plot_facet_map(ad_prediction_list[[i]], "est_exp",
      raster_limits = c(0, max_raster),
      # legend_position = "none"
      transform_col = fourth_root_power
    ) +
      labs(fill = "kg/ha") +
      ggtitle(paste0("", species, " mature biomass \n(max = ", max_adult, " kg/ha)"))
    p_adult[[i]]
  }
  print(p_adult)
}
```

```{r, message=FALSE, warning=FALSE}
# try(out_mature <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")))
rm(out_mature)
if (!exists("out_mature")) {
  if (exists("adult_predictions")) {
    predictions <- adult_predictions
  } else {
    try(out_total <- readRDS(paste0(
      "data/", spp, "/trends-", spp, covs, "-", ssid_string, "-total-biomass-priors-", priors, ".rds"
    )))
  }

  if (!exists("out_total")) {
    if (1 %in% model_ssid) {
      out_mature1 <- make_trend_data(predictions,
        ssid = c(1),
        start_time = 2003,
        skip_time = 2004,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = c(1, 1, 1, 1, 2, 2, 2, 2),
        variable_names = "est_exp"
      )
    } else {
      out_mature1 <- NULL
    }

    if (3 %in% model_ssid) {
      out_mature3 <- make_trend_data(predictions,
        ssid = c(3),
        start_time = 2005,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = c(1, 1, 1, 2, 2, 2, 2),
        variable_names = "est_exp"
      )
    } else {
      out_mature3 <- NULL
    }

    if (4 %in% model_ssid) {
      out_mature4 <- make_trend_data(predictions,
        ssid = c(4),
        start_time = 2004,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = c(1, 1, 1, 1, 2, 2, 2, 2),
        variable_names = "est_exp"
      )
    } else {
      out_mature4 <- NULL
    }

    if (16 %in% model_ssid) {
      out_mature16 <- make_trend_data(predictions,
        ssid = c(16),
        start_time = 2004,
        skip_time = 2007,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = c(1, 1, 1, 2, 2, 2, 2),
        variable_names = "est_exp"
      )
    } else {
      out_mature16 <- NULL
    }

@@

    saveRDS(out_mature, file = paste0(
      "data/", spp, "/trends-", spp, covs, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds"
    ))

    if (!exists("adult_predictions")) {
      out_total <- out_mature
      saveRDS(out_total, file = paste0(
        "data/", spp, "/trends-", spp, covs, "-", ssid_string, "-total-biomass-priors-", priors, ".rds"
      ))
    }
  }
}
```

### Recent biomass with trawl footprint
```{r }
if (exists("out_mature")) {
  trawl_footprint <- sf::st_read(dsn = "data/trawl-footprint", layer = "Trawl_footprint")
  proj.to <- "+proj=utm +zone=9 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"
  trawl_footprint <- sf::st_transform(trawl_footprint, crs = proj.to)

  current_biomass <- plot_vocc(out_mature,
    fill_col = "est_exp_e",
    fill_label = "Mean mature \nbiomass (kg/ha)\npost-2010",
    raster_alpha = 1,
    viridis_option = "C",
    vec_aes = NULL,
    transform_col = fourth_root_power
  )
  current_biomass <- current_biomass +
    geom_sf(data = trawl_footprint, aes(geometry = geometry / 1000), colour = "gray", fill = "gray", alpha = 0.2, lwd = 0.25, inherit.aes = FALSE) +
    coord_sf(xlim = c(180, 820), ylim = c(5350, 6050), datum = NULL) +
    xlab("") +
    ylab("") +
    theme(panel.grid.major = element_line(colour = "transparent")) +
    ggtitle(paste("Predicted", species, "density with trawl footprint overlain"))
  current_biomass
}
```

@@

    if (1 %in% model_ssid) {
      out_im1 <- make_trend_data(predictions,
        ssid = c(1),
        start_time = 2003,
        skip_time = 2004,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = c(1, 1, 1, 1, 2, 2, 2, 2),
        variable_names = "est_exp"
      )
    } else {
      out_im1 <- NULL
    }

    if (3 %in% model_ssid) {
      out_im3 <- make_trend_data(predictions,
        ssid = c(3),
        start_time = 2005,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = c(1, 1, 1, 2, 2, 2, 2),
        variable_names = "est_exp"
      )
    } else {
      out_im3 <- NULL
    }

    if (4 %in% model_ssid) {
      out_im4 <- make_trend_data(predictions,
        ssid = c(4),
        start_time = 2004,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = c(1, 1, 1, 1, 2, 2, 2, 2),
        variable_names = "est_exp"
      )
    } else {
      out_im4 <- NULL
    }

    if (16 %in% model_ssid) {
      out_im16 <- make_trend_data(predictions,
        ssid = c(16),
        start_time = 2004,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = c(1, 1, 1, 1, 2, 2, 2, 2),
        variable_names = "est_exp"
      )
    } else {
      out_im16 <- NULL
    }

@@

Save spatiotemporal biomass plots
```{r results = "hide"}
biomass_plots <- c(p_adult, p_imm)

>>>>

---
title: "`r paste(params$species)` biomass trends and figures"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    species: "Petrale Sole"
    region: "All synoptic surveys"
    start_year: 2008
    bind_predictions: FALSE
    covs: "-trawled-ssid" # covariate string used to name model rds 
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set( fig.width = 9, fig.height = 6.5, #fig.width=11, fig.height=8.5, 
                      #fig.path=paste0("figs/", spp, "/"),
                      echo=FALSE, warning=FALSE, message=FALSE)
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)
```

```{r params}
species <- params$species
region <- params$region
start_year <- params$start_year

bind_predictions <- params$bind_predictions
covs <- params$covs
priors <- FALSE

@@

if (region == "All synoptic surveys") {
  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
  model_ssid <- c(1, 3, 4, 16)
  ssid_string <- paste0(model_ssid, collapse = "n")
}

if (region == "both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
}
```

@@

if (region != "All synoptic surveys") {
  try(adult_predictions <- readRDS(paste0("data/", spp,
    "/predictions-", spp, covs, "-1n3n4n16-mature-biomass-prior-", priors, ".rds"
  )) %>% filter(ssid %in% model_ssid))
  try(imm_predictions <- readRDS(paste0("data/", spp,
    "/predictions-", spp, covs, "-1n3n4n16-imm-biomass-prior-", priors, ".rds"
  )) %>% filter(ssid %in% model_ssid))

@@

    try(predictions <- readRDS(paste0("data/", spp,
      "/predictions-", spp, covs, "-1n3n4n16-total-biomass-prior-", priors, ".rds"
    )) %>% filter(ssid %in% model_ssid))

@@

Caluclate variables of interest from predictions
```{r}
if (exists("adult_predictions")) {
  # to convert from kg/m2 to kg/hectare multiply by 10000
  adult_predictions$est_exp <- exp(adult_predictions$est) * 10000
  imm_predictions$est_exp <- exp(imm_predictions$est) * 10000

@@

  adult_predictions <- filter(adult_predictions, year >= start_year)
  imm_predictions <- filter(imm_predictions, year >= start_year)

@@

  model_ssid <- unique(adult_predictions$ssid)
} else {
  predictions <- filter(predictions, year >= start_year)

  predictions$est_exp <- exp(predictions$est) * 10000
  predictions$total_bio <- exp(predictions$est) * 10000
  max_raster <- quantile(predictions$total_bio, .999)
  #max_bio <- round(max(predictions$total_bio))
  max_bio <- signif(max(predictions$est_exp), digits = 2)

@@

### Split by survey (SSID)
```{r}
if (exists("adult_predictions")) predictions <- adult_predictions

  if (1 %in% model_ssid) {
    adult_predictions1 <- predictions %>% filter(ssid == 1)
  } else {
    adult_predictions1 <- NULL
  }

  if (3 %in% model_ssid) {
    adult_predictions3 <- predictions %>%
      filter(ssid != 16) %>%
      filter(ssid != 4)
  } else {
    adult_predictions3 <- NULL
  }

  if (4 %in% model_ssid) {
    adult_predictions4 <- predictions %>% filter(ssid == 4)
  } else {
    adult_predictions4 <- NULL
  }

  if (16 %in% model_ssid) {
    if (any(predictions$ssid == 16)) {
      adult_predictions16 <- predictions %>% filter(ssid == 16)
    } else {
      adult_predictions16 <- NULL
    }
  } else {
    adult_predictions16 <- NULL
  }

@@

  p_adult <- list()
  if(exists("adult_predictions")) {
  for (i in seq_len(length(model_ssid))) {
    p_adult[[i]] <- plot_facet_map(ad_prediction_list[[i]], "est_exp",
      raster_limits = c(0, max_raster),
      # legend_position = "none"
      transform_col = fourth_root_power
    ) +
      labs(fill = "kg/ha") +
      ggtitle(paste0("", species, " mature biomass \n(max = ", max_adult, " kg/ha)"))
    p_adult[[i]]
  }
  } else {
    for (i in seq_len(length(model_ssid))) {
    p_adult[[i]] <- plot_facet_map(ad_prediction_list[[i]], "est_exp",
      raster_limits = c(0, max_raster),
      # legend_position = "none"
      transform_col = fourth_root_power
    ) +
      labs(fill = "kg/ha") +
      ggtitle(paste0("", species, " total biomass \n(max = ", max_bio, " kg/ha)"))
    p_adult[[i]]  
    }
  }
  print(p_adult)
```

```{r, message=FALSE, warning=FALSE}
rm(out_mature)
rm(out_total)
# try(out_mature <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")))

if (!exists("out_mature")) {
  if (exists("adult_predictions")) {
    predictions <- adult_predictions
  } else {
   try(out_total <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-total-biomass-priors-", priors, ".rds")))
  }

  if (!exists("out_total")) {
    if (1 %in% model_ssid) {

      .predictions <- filter(predictions, ssid == 1)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_mature1 <- make_trend_data(predictions,
        ssid = c(1),
        start_time = start_year,
        skip_time = 2004,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_mature1 <- NULL
    }

    if (3 %in% model_ssid) {
      .predictions <- filter(predictions, ssid == 3)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_mature3 <- make_trend_data(predictions,
        ssid = c(3),
        start_time = start_year,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_mature3 <- NULL
    }

    if (4 %in% model_ssid) {

      .predictions <- filter(predictions, ssid == 4)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_mature4 <- make_trend_data(predictions,
        ssid = c(4),
        start_time = start_year,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_mature4 <- NULL
    }

    if (16 %in% model_ssid) {

      .predictions <- filter(predictions, ssid == 16)
      #unique(.predictions$year)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_mature16 <- make_trend_data(predictions,
        ssid = c(16),
        start_time = start_year,
        skip_time = 2007,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_mature16 <- NULL
    }

@@

    if (exists("adult_predictions")) {
            saveRDS(out_mature, file = paste0("data/", spp, "/trends-", 
        spp, covs, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds"))

         } else {
      out_total <- out_mature
      rm(out_mature)
      saveRDS(out_total, file = paste0("data/", spp, "/trends-", 
        spp, covs, "-", ssid_string, "-total-biomass-priors-", priors, ".rds"))
    }
  }
}
```

### Recent biomass with trawl footprint
```{r }
if (!exists("out_mature")) {out_recent <- out_total} else {out_recent <- out_mature}

  trawl_footprint <- sf::st_read(dsn = "data/trawl-footprint", layer = "Trawl_footprint")
  proj.to <- "+proj=utm +zone=9 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"
  trawl_footprint <- sf::st_transform(trawl_footprint, crs = proj.to)

  current_biomass <- plot_vocc(out_recent,
    fill_col = "est_exp_e",
    fill_label = "Mean\nbiomass (kg/ha)\npost-2010",
    raster_alpha = 1,
    viridis_option = "C",
    vec_aes = NULL,
    transform_col = fourth_root_power
  )
  current_biomass <- current_biomass +
    geom_sf(data = trawl_footprint, aes(geometry = geometry / 1000), colour = "gray", fill = "gray", alpha = 0.2, lwd = 0.25, inherit.aes = FALSE) +
    coord_sf(xlim = c(180, 820), ylim = c(5350, 6050), datum = NULL) +
    xlab("") +
    ylab("") +
    theme(panel.grid.major = element_line(colour = "transparent")) +
    ggtitle(paste("Predicted", species, "density with trawl footprint overlain"))
  current_biomass

```

@@

    if (1 %in% model_ssid) {

      .predictions <- filter(predictions, ssid == 1)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
      out_im1 <- make_trend_data(predictions,
        ssid = c(1),
        start_time = start_year,
        skip_time = 2004,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_im1 <- NULL
    }

    if (3 %in% model_ssid) {
      .predictions <- filter(predictions, ssid == 3)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_im3 <- make_trend_data(predictions,
        ssid = c(3),
        start_time = start_year,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_im3 <- NULL
    }

    if (4 %in% model_ssid) {
      .predictions <- filter(predictions, ssid == 4)
      unique(.predictions$year)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_im4 <- make_trend_data(predictions,
        ssid = c(4),
        start_time = start_year,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_im4 <- NULL
    }

    if (16 %in% model_ssid) {
      .predictions <- filter(predictions, ssid == 16)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_im16 <- make_trend_data(predictions,
        ssid = c(16),
        start_time = start_year,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_im16 <- NULL
    }

@@

Save spatiotemporal biomass plots
```{r results = "hide"}
if(exists("adult_predictions")) { biomass_plots <- c(p_adult, p_imm) } else {
  biomass_plots <- c(p_adult)
}
